---
title: " Shiny Application and Reproducible Pitch"
author: "Nikhil Chandra"
date: "September 23 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

* This pitch and the shiny application was built for the "Developing Data Products" assignment.
* This application can calculate BMI once you enter weight and Height.

## Overview of Application

* This application can calculate BMI once you enter weight and Height.
You need to enter weight in Kg

* Enter Height in meters.

* Enter the date 

* Click Submit.

## Result of Application

* After providing the above input, the application will diplay the below result.

* Height

* Weight

* Date

* BMI



<br /><br />

* The developed shiny app for this assignment is avalilable in this presentation and at: https://nikhilchandra.shinyapps.io/ShiinyApplication-BMI/

<br /><br />

* The source codes of ui.R and server.R and also the presentation are available on the GitHub repo: 
https://github.com/nikhil-chandra/datasciencecoursera/tree/master/shiny%20Application





## Shiny App: Example

```{r}
knitr::include_app("
https://nikhilchandra.shinyapps.io/ShiinyApplication-BMI/", height = "450px")
```
The developed shiny app for this assignment is also avalilable at: 
https://nikhilchandra.shinyapps.io/ShiinyApplication-BMI/