---
title: "Peer-Assignment:R Markdown,Presentation & Plotly"
author: "Nikhil Chandra"
date: "September 23 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(comment = "")
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
```

## Instructions

Create a web page presentation using R Markdown that features a plot created with Plotly. Host your webpage on either GitHub Pages, RPubs, or NeoCities. Your webpage must contain the date that you created the document, and it must contain a plot created with Plotly. We would love to see you show off your creativity!

## Load and Prepare Data
```{r, echo = TRUE}
library(plotly)
data <- diamonds[sample(nrow(diamonds), 2500), 
                 c("carat", "price", "clarity", "depth")]
summary(data)
```

## Create 2D and 3D Scatter Plots
```{r, echo = TRUE}
p1 <- plot_ly(data, x = ~carat, y = ~price, color = ~carat,
        size = ~carat, text = ~paste("Clarity: ", clarity))

p2 <- plot_ly(data, x = ~carat, y = ~price, z = ~depth,
        color = ~carat, size = ~carat, 
        text = ~paste("Clarity: ", clarity)) 
```

## Show 2D Scatter Plot

```{r}
ggplotly(p1)
```

## Show 3D Scatter Plot

```{r}
ggplotly(p2)
```
