---
title: "Statistical Inference - Peer Grading"
author: "Nikhil Chandra"
date: "October 2, 2018"
output:
  html_document:
    df_print: paged
---


## Effectiveness of Vitamin C on Tooth Growth in Guinea Pigs

## Overview
In this project, we are going to analyze the ToothGrowth data in the dataset .

        1. We are going to load the ToothGrowth data and do some basic exploratory data analysis.
        2. Then we will provide a summary for the data.
        3. Usage of hypothesis or confidence intervals tests to compare tooth growth by supplement and dose. 
        4. Provide assumptions and the Conclusion.



```{r}
        library(dplyr)
        library(ggplot2)
        library(datasets)
        library(gridExtra)
        library(GGally)
```

## 1. Load the ToothGrowth data - Perform basic Analysis.

### Load dataset - ToothGrowth
```{r}
        data(ToothGrowth)
        # to keep  naming format with lowercase 1st letter and uppercase subsequent letters.
        toothGrowth <- ToothGrowth
        # convert to factor for plotting
        toothGrowth$dose <- as.factor(toothGrowth$dose)
```

### Some Exploratory Analysis.
```{r}
        str(toothGrowth)
        head(toothGrowth)
        tail(toothGrowth)
        summary(toothGrowth)
```

### Number of Rows and Columns.
```{r}
        dim(toothGrowth)
```

### Sample Size n
```{r}
        sample_size <- length(toothGrowth$len)
        sample_size
```

### Mean group by dose 
```{r}
        ## X bar - Mean
        mean_groups <- aggregate(toothGrowth$len,list(toothGrowth$supp,toothGrowth$dose),mean)
        mean_groups
```

### Standard Deviation - group by dose 
```{r}
        ## s - standard Deviation
        sd_group <- aggregate(toothGrowth$len,list(toothGrowth$supp,toothGrowth$dose),sd)
        sd_group
```

# 2. Summary of the data.

```{r}
        summary(toothGrowth)
```

### Table
```{r}
        table(toothGrowth$supp, toothGrowth$dose)
```

### Plots
```{r echo=FALSE}
p1 <- ggplot(data=toothGrowth, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~supp) 

p2 <- ggplot(data=toothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~dose) 

p3 <- ggplot(data=toothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot()

p4 <- ggplot(data=toothGrowth, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot()

grid.arrange(p1, p4, p2, p3, ncol = 2, nrow=2, top="Boxplots - ToothGrowth Data by Vit C Dose and Supplement")
```

# 3. Usage of hypothesis or confidence intervals tests to compare tooth growth by supplement and dose. 

Perform - Analysis of Variance
```{r}
anova.out <- aov(len ~ supp * dose, data=toothGrowth)
summary(anova.out)
```

The Analysis of Variance results show interaction between the dosage and length. (F(1,54)=15.572;p<0.01). 
We can alsp see that the supplement type (supp) (F(2,54)=92.000;p<0.01) effects length. 
We also see some amount of  interaction between the combination of supplement type  and dosage compared to the length  (F(2,54)=4.107;p<0.05).

```{r}
TukeyHSD(anova.out)
confint(anova.out)
print(model.tables(anova.out,"means"),digits=3)
```

The Analysis of the Tukey HSD  analysis shows some significant differences between each of the groups in supp and dose.

# 4. Assumptions and Conclusion. 

#### With the above data we can  conclude that the data indicates that both the dosage and  supplement have  effects on the length growth of teeth of guinea pig .
 Supplement type has also shown influence. OJ has a greater average teeth growth with the combination of dosages at 0.5 and 1 than for the VC supplement. 
But, at dose level 2, there seems to be no major significant effect (similar means and confidence intervals) between the VC supplement and the OJ.

The above conclusions are based on the following assumptions:

a. The distribution of the means is approximately normal.
b. Dosage and supplement are randomly assigned.
